Universal Targeted Attacks for Siamese \\ Visual Tracking

### 2021-04-22 博士/论文提交/adversarial_attack/TCSVT/latex/bare_jrnl.tex
初稿完成
完善

% 题目
\hyphenation{op-tical net-works semi-conduc-tor}
\begin{document}
\title{Universal Targeted Attacks for Siamese \\ Visual Tracking}
\author{
  Zhenbang Li
  \thanks{This work is supported by the National Key R\&D Program of China (No. 2018AAA0102802, No. 2018AAA0102803, No. 2018AAA0102800), the NSFC-General Technology Collaborative Fund for Basic Research (Grant No. U1636218), the Natural Science Foundation of China (Grant No. 61751212, 61721004, 61772225, 61972394), the Key Research Program of Frontier Sciences, CAS (Grant No. QYZDJ-SSW-JSC040), and the National Natural Science Foundation of Guangdong (No. 2018B030311046). \textit{(Corresponding author: Jin Gao.)}}

  \thanks{Z. Li is with the National Laboratory of Pattern Recognition, Institute of Automation, Chinese Academy of Sciences, Beijing, 100190, China, and also with School of Artificial Intelligence, University of Chinese Academy of Sciences, Beijing, 100190, China (e-mail: zhenbang.li@nlpr.ia.ac.cn; bli@nlpr.ia.ac.cn; jin.gao@nlpr.ia.ac.cn).}

  \iffalse
  \thanks{L. Li is with the Brain Science Center, Beijing Institute of Basic Medical Sciences, Beijing, China (e-mail: liang.li.brain@aliyun.com).}
  \thanks{W. Hu is with CAS Center for Excellence in Brain Science and Intelligence Technology, National Laboratory of Pattern Recognition, Institute of Automation, Chinese Academy of Sciences, Beijing, 100190, China, and also with School of Artificial Intelligence, University of Chinese Academy of Sciences, Beijing, 100190, China (e-mail: wmhu@nlpr.ia.ac.cn).}
  \fi
}

@@

% 摘要
\begin{abstract}
Siamese trackers are shown to be vulnerable to adversarial attacks recently. However, the existing attack methods craft the perturbations for each video independently, which comes at a non-negligible computational cost. The question is what if we can not get access to the limited computational resources in the real-world online-tracking phase.

In this paper, we show the existence of universal perturbations that can enable the targeted attack, e.g., forcing a tracker to follow the ground-truth trajectory with specified offsets, to be universal and free from inference in a network. Specifically, we attack a tracker by adding a universal imperceptible perturbation to the template image and pasting a \textit{fake target}, i.e., a small universal adversarial patch, into the search images adhering to the predefined trajectory, so that the tracker outputs the location and size of the \textit{fake target} instead of the real target. Our approach allows perturbing a novel video to come at no additional cost except the mere addition and pasting operations -- and not require gradient optimisation or network inference. Experimental results on several datasets demonstrate that our approach can effectively fool the Siamese trackers in a targeted attack manner: the 16-by-16 adversarial patch can make the AO score drop dramatically to 0.12 and its area only accounts for 0.39\% of the search image area.
%The attack performance is state-of-the-art: drop the presion score of OTB2015 from 0.861 to 0.048, and 0.928 for the fake trajectory. 
The proposed method can also perform black box attack on different tracking architectures.
We will make our code publicly available.
\end{abstract}

% 关键字
\begin{IEEEkeywords}
Visual tracking, adversarial attack, siamese networks.
\end{IEEEkeywords}
\IEEEpeerreviewmaketitle

@@

Recently, the robustness of Siamese trackers have attracted much attention in the sense of testing their vulnerability to adversarial attacks, and the focus has been directed towards more efficient and low-cost attacks \cite{TTP,FAN,SPARK}. Despite their success, these attack methods are not suitable for attacking the real-world online-tracking systems built upon the small edge platforms with limited computational resources. The reason is that they still need to craft the perturbations for each video independently based on either iterative optimisation or adversarial network inference, for which the adequacy of computational resources may be not ensured by the computationally intensive tracking systems.

Universal adversarial perturbations (UAPs) proposed in \cite{UAP} can fool most images from a data distribution in an image-agnostic manner. Being universal, UAPs can be conveniently exploited to perturb unseen data on-the-fly without extra computations. Therefore, UAPs are particularly useful when attacking real-time applications deployed on platforms with limited computational resources. However, no existing work has touched the topic of attacking the Siamese trackers using UAPs, because it is hard to apply existing UAPs to attack Siamese trackers directly. The main reason lies in the fact that, (a) most UAPs are designed for typical neural networks with one image as input while Siamese networks accept both the template and search images, and (b) the goal of existing UAP methods is to disturb unary or binary model outputs for single instance while we need to use universal perturbations to mislead Siamese trackers to follow a specified trajectory.

@@

@@

\textit{White box attacks v.s. Black box attacks} In the white box attack setting \cite{meng2019white}, the adversary has full knowledge of the model including model type, model architecture and values of all parameters and trainable weights. In the black box setting \cite{cheng2018query,li2019nattack,papernot2017practical,li2020projection}, the adversary has limited or no knowledge about the model under attack \cite{kurakin2018adversarial}. In this paper, we focus on the white box attack for siamese trackers. However, although we train perturbations using the parameters of the SiamFC++ network, we surprisely find that the trained perturbations can perform black attack on other trackers such as SiamRPN and SiamRPN++.

\textit{Non-universal attacks v.s. Universal attacks} In the setting of non-universal attacks \cite{dai2018adversarial,li2018second,lin2017tactics}, the adversary has to generate one perturbation of every new datapoint, which is time consuming. In the setting of the universal attacks \cite{khrulkov2018art,mopuri2018nag,zhang2020understanding,mopuri2018generalizable,chen2018shapeshifter}, only one perturbation is generated and used for every datapoint of the dataset. In this paper, we focus on generating universal perturbations for siamese trackers to perform efficient attack.

@@

Let $V=\{I_i\}_1^T$ denote the frames of a video sequence of length $T$. $B^{gt}=\{b^{gt}_i\}_1^T$ is used to represent the target's ground-truth position in each frame. The visual object tracking aims to predict the position $B^{pred}=\{b^{pred}_i\}_1^T$ of this target in the subsequent frames given its initial state. In SiamFC++, the tracker first transforms the paired reference frame $I_1$ and annotation $b_1^{gt}$ to get an template image $\textbf z$, and transforms the search frame $I_i$ to get the search image $\textbf x_i$ centered at the position estimated in the previous frame. At each time-step, the template image $\textbf z$ and the search image $\textbf x_i$ are first passed individually through a shared backbone network and then fused using a channel-by-channel mutual correlation operation:
\begin{equation}
  f_{i}(\mathbf{z}, \mathbf{x})=\psi_{i}(\phi(\mathbf{z})) \star \psi_{i}(\phi(\mathbf{x})), i \in\{\mathrm{cls}, \text {reg}\}
\end{equation}
where $\star$ denotes the channel-by-channel correlation operation, $\phi(\cdot)$ denotes the feature extractor of the twin network, $\psi_i(\cdot)$ denotes the layer specific to the classification or regression task, and $i$ denotes the specific task type ($\mathrm{cls}$ denotes the classification task and $\mathrm{reg}$ denotes the regression task). $\psi_{\mathrm{cls}}$ and $\psi_{\mathrm{reg}}$ are both designed as two convolutional layers for adapting generic features to the feature space specific to the classification/regression task. $\psi_{\mathrm{cls}}$ and $\psi_{\mathrm{reg}}$ extracted features have the same size. The fused features then act as input to a head network, which predicts a classification map $\textbf{C}$, a bounding box regression map $\textbf{R}$, and a quality assessment map $\textbf{Q}$ in an anchor-free manner. In short, $\textbf C$ encodes the probability of each spatial position to contain the target, $\textbf R$ regresses the bounding box of the target, and $\textbf Q$ predicts the target state estimation quality. The final bounding box is then generated according to $\textbf{C}$, $\textbf{R}$ and $\textbf{Q}$.

@@

Our attack algorithm is based on the imperceptible perturbation genration method \cite{FGSM} and adversarial patch generation method \cite{patch}. However, we do not simply apply either of above methods to trackers. In this paper, we demostrate that the imperceptible perturbations and the adversarial patch are indispensable for attacking trackers and can be jointly optimized in and end-to-end training manner. On the one hand, the small patch applied to the search image works as a fake target, which tells the tracker where the target is. It is crucial to design a strategy for the tracker to follow a predefined trajectory. For example, the TTP method \cite{TTP} pre-computes perturbations corresponding diverse directions and use them to force the tracker to follow the target trajectory. However, this method is sub-optimal because the tracker only predicts the approximation of the specified trajectory. As shown in experiments, the targeted attack performance of TTP is not as good as ours. On the other hand, the imperceptible perturbation applied to the template image is optimized to find the fake target, cooling down hot regions where the real target exists on the heatmaps and increasing the responses at the position of the fake target (see Figure \ref{fig:1}).

@@

In this subsection, we show how to train the video-agnostic perturbations $(\delta, p)$ for Siamese trackers (Figure \ref{fig:vis_perturbations}). At beginning, each element in $\delta$ is initialized to 0. $p$ is initialized with a normal distribution, with a mean of 127 and a standard deviation of 1. During the $k$-th iteration of training, a video $V=\{I_i\}_1^T$ is randomly selected from the training dataset $\mathcal V$. Assuming the template perturbation at the $k$-th iteration is $\delta_k \in \mathbb{R}^{127\times 127 \times 3}$, and the adversarial patch is $p_k \in \mathbb{R}^{128\times 128\times 3}$. We first randomly pick paired frames $I_t, I_s$ from $V$. The clean template image $\textbf z\in \mathbb{R}^{127\times 127 \times 3}$ is generated according to $I_t$ and $b^{gt}_t$, and the perturbed template image is:
\begin{equation}
\tilde {\textbf z} = \textbf z + \delta_k.
\end{equation}
Similarly, the clean search image $\textbf x \in \mathbb{R}^{303\times 303 \times 3}$ is generated according to $I_s$ and $b^{gt}_s$.
As mentioned before, the patch is regarded as a \textit{fake target} and pasted into the search images. We force the center position of the \textit{fake target} to near the center position of the real target within a shift range of 64 pixels, where shift is defined as the maximum range of translation generated from a uniform distribution.
The width/height of the \textit{fake target} is randomly selected between 32 pixels and 128 pixels.
The perturbed search image is generated as follows:
\begin{equation}
\tilde{\textbf x} = A(\textbf x, p_k, (l^x, l^y), (w, h)),
\end{equation}
where $(l^x, l^y)$ and $(w, h)$ represent the position and size of the \textit{fake target} with respect to the search image, respectively. $A$ is a patch application operator \cite{patch} which first resizes the patch $p_k \in \mathbb{R}^{128\times 128\times 3}$ to $\hat{p}_k \in \mathbb{R}^{w\times h\times 3}$, and then pastes the resized patch $\hat{p}_k$ into the search image $\textbf x$ at location $(l^x,l^y)$. Subsequently, the SiamFC++ tracker $\phi(\cdot)$ takes $\tilde {\textbf x}$ and $\tilde{\textbf  z}$ as input and makes predictions as follows:
\begin{equation}
\textbf{C, R, Q} = \phi(\tilde {\textbf x}, \tilde{\textbf z}).
\end{equation}

\begin{algorithm}[tb]
  \caption{Training Process}
  \label{alg:algorithm}
  \textbf{Input}: Training dataset $\mathcal{V}$, Siamese tracker $\phi$, and max iteration number $N$.\\
  \textbf{Output}: Imperceptible perturbation $\delta$, and adversarial patch $p$.
  \begin{algorithmic}[1] %[1] enables line numbers
  \STATE Let $k = 0$.
  \WHILE{$k < N$}
  \STATE Randomly pick a video $V\in \mathcal{V}$. The corresponding ground truth is $B^{gt}=\{b^{gt}_i\}^T_1$.
  \STATE Randomly pick paired frames $I_t, I_s$ from $V$.
  \STATE Generate template image $\textbf{z}$ according to $I_t$ and $b^{gt}_t$.
  \STATE $\tilde{\textbf{z}} = \textbf{z} + \delta_k.$
  \STATE Generate search image $\textbf{x}$ according to $I_s$ and $b^{gt}_s$.
  \STATE Calculate the \textit{fake target} position $\{l^x, l^y, w, h\}$ with respect to the search image.
  \STATE $\tilde{\textbf x} = A(\textbf x, p_k, (l^x, l^y), (w, h)).$
  \STATE $\textbf{C, R, Q} = \phi(\tilde {\textbf x}, \tilde{\textbf z}).$
  \STATE Generate fake labels $\textbf{C}^*,\textbf{R}^*,\textbf{Q}^*$ using $\{l^x, l^y, w, h\}$.
  \STATE Calculate loss $L(\textbf{C, R, Q}, \textbf{C}^*, \textbf{R}^*, \textbf{Q}^*)$ using Equ. \ref{eq:loss}.
  \STATE $\delta_{k+1} = \delta_{k} - \epsilon_1 \cdot \text{sign}(\nabla_{\delta_k}L).$
  \STATE $p_{k+1} = p_{k} - \epsilon_2 \cdot \text{sign}(\nabla_{p_k}L).$
  \STATE $k = k + 1.$
  \ENDWHILE
  \STATE \textbf{return} $\delta_N, p_N.$
  \end{algorithmic}
  \label{alg}
\end{algorithm}

\textit{Generating Fake Labels} The fake lables are composes of three parts: fake classification label $\textbf{C}^*$, fake regression label $\textbf{R}^*$ and fake quality estimation label $\textbf{Q}^*$.
For classification, location $(x,y)$ on feature map $\psi_{\mathrm{cls}}$ is considered as a positive sample (i.e., $\textbf{C}^*_{x,y} = 1$) if its corresponding location $\left(\left\lfloor\frac{s}{2}\right\rfloor+x s,\left\lfloor\frac{s}{2}\right\rfloor+y s\right)$ on the input image falls into the fake bounding box, and vice versa are considered as negative samples (i.e., $\textbf{C}^*_{x,y} = 0$). $s=8$ denotes the total step length of the feature extraction network.

In the regression branch, the last convolution layer predicts the distance of position $\left(\left\lfloor\frac{s}{2}\right\rfloor+x s,\left\lfloor\frac{s}{2}\right\rfloor+y s\right)$ to the four edges of the fake target labeled box, denoted as the four-dimensional vector $\boldsymbol{t}^{*}=\left(l^{*}, t^{*}, r^{*}, b^{*}\right)$. Thus, the regression label of position $(x,y)$ can be expressed as:

\begin{equation}
  \begin{array}{ll}
  l^{*}=\left(\left\lfloor\frac{s}{2}\right\rfloor+x s\right)-x_{0}, \quad t^{*}=\left(\left\lfloor\frac{s}{2}\right\rfloor+y s\right)-y_{0} \\
  r^{*}=x_{1}-\left(\left\lfloor\frac{s}{2}\right\rfloor+x s\right), \quad b^{*}=y_{1}-\left(\left\lfloor\frac{s}{2}\right\rfloor+y s\right)
  \end{array}
\end{equation}
where $(x_0, y_0)$ and $(x_1, y_1)$ denote the coordinates of the upper-left and lower-right corners of the fake target labeling box $B^*$, respectively.

SiamFC++ assumes that feature pixels near the center of the target will have higher importance than other pixels. Following the design of SiamFC++, we use a $1 \times 1$ convolutional layer for quality assessment, i.e., to learn the intersection over union (IoU) score of the predicted border to the fake target labeled box:

\begin{equation}
  \mathrm{IoU}^{*}=\frac{\text { Intersection }\left(B, B^{*}\right)}{\operatorname{Union}\left(B, B^{*}\right)}
\end{equation}

@@

\textit{Ablation Study: Influence of Training Iterations} Table \ref{tab:iter} and Figure \ref{fig:iter} demonstrate the effect of the number of training iterations of the algorithm on the effectiveness of the attack. It can be observed that as the number of iterations increases, the AO value with respect to the false trajectory increases significantly, while the AO value with respect to the real trajectory decreases significantly. This demonstrates the effectiveness of the end-to-end training process proposed in this chapter.
After about 30,000 training iterations, the resulting perturbation prevents the tracker from tracking most of the targets in GOT-Val, and the AO with respect to the true trajectory decreases from 0.760 to 0.035. We note that the AO decreases significantly faster at the beginning of the training cycle (training iterations less than 2048) than later. This demonstrates the fast convergence capability of our end-to-end training scheme.

To verify whether the perturbations generated for the template images are difficult to perceive, we evaluate and analyze the quality of the template images after the perturbations. The quality of the perturbed template images is evaluated by the metric SSIM. As shown in Table \ref{tab:iter}, the SSIM value of the perturbed template image gradually decreases as the training proceeds, which indicates that the difference between the perturbed template image and the original template image gradually increases due to the learned adversarial perturbation being added to the clean template image. The SSIM value ranges from 0 to 1. If the SSIM is close to 1, the difference between the perturbed template image and the original template image is small. If SSIM is closer to 0, it means that the difference between the perturbed template image and the original template image is large, which means that the perturbation generated by the template image is easier to perceive. The SSIM value after convergence of this algorithm is 0.88, which means that the difference between the perturbed template image and the original template image is small, that is, the perturbation generated by the template image is harder to perceive (see Figure \ref{fig:1}).

@@

% 作者
%\vfill  % 距离更远了
\enlargethispage{-5in}  % 距离近了些
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{images/zhbli.jpg}}]{Zhenbang Li}
received the B.S. degree in computer science and technology from Beijing Institute of Technology, Beijing, China, in 2016. Currently, he is a Ph.D. student in Institute of Automation, Chinese Academy of Sciences, Beijing, China. His research interests include object tracking and deep learning.
\end{IEEEbiography}

>>>>

% 题目
\hyphenation{op-tical net-works semi-conduc-tor}
\begin{document}
\title{Universal Targeted Attacks for Siamese \\ Visual Tracking}
\author{
  Zhenbang Li, Jin Gao, Yaya Shi, Bing Li, Weiming Hu
  \thanks{This work is supported by the national key R\&D program of china (No. 2018AAA0102802, No. 2018AAA0102803, No. 2018AAA0102800), the Natural Science Foundation of China (Grant No. 62036011, 61721004, 61772225), the Key Research Program of Frontier Sciences, CAS, Grant No. QYZDJ-SSW-JSC040, and the National Natural Science Foundation of Guangdong (No. 2018B030311046). \textit{(Corresponding author: Jin Gao.)}}

  \thanks{Z. Li, J. Gao and B. Li are with the National Laboratory of Pattern Recognition, Institute of Automation, Chinese Academy of Sciences, Beijing, 100190, China, and also with School of Artificial Intelligence, University of Chinese Academy of Sciences, Beijing, 100190, China (e-mail: zhenbang.li@nlpr.ia.ac.cn; jin.gao@nlpr.ia.ac.cn; bli@nlpr.ia.ac.cn).}
  \thanks{Y. Shi is with the University of Science and Technology of China, Anhui, 230026, China (e-mail: shiyaya@mail.ustc.edu.cn).}
  \thanks{W. Hu is with CAS Center for Excellence in Brain Science and Intelligence Technology, National Laboratory of Pattern Recognition, Institute of Automation, Chinese Academy of Sciences, Beijing, 100190, China, and also with School of Artificial Intelligence, University of Chinese Academy of Sciences, Beijing, 100190, China (e-mail: wmhu@nlpr.ia.ac.cn).}
}

@@

% 摘要
\begin{abstract}
Siamese trackers are shown to be vulnerable to adversarial attacks recently. However, the existing attack methods craft the perturbations for each video independently, which comes at a non-negligible computational cost. In this paper, we show the existence of universal perturbations that can enable the targeted attack, e.g., forcing a tracker to follow the ground-truth trajectory with specified offsets, to be video-agnostic and free from inference in a network. Specifically, we attack a tracker by adding a universal imperceptible perturbation to the template image and pasting a \textit{fake target}, i.e., a small universal adversarial patch, into the search images adhering to the predefined trajectory, so that the tracker outputs the location and size of the \textit{fake target} instead of the real target. Our approach allows perturbing a novel video to come at no additional cost except the mere addition and pasting operations -- and not require gradient optimisation or network inference. Experimental results on several datasets demonstrate that our approach can effectively fool the Siamese trackers in a targeted attack manner: the 16-by-16 adversarial patch can make the AO score drop dramatically to 0.12 and its area only accounts for 0.39\% of the search image area.
%The attack performance is state-of-the-art: drop the presion score of OTB2015 from 0.861 to 0.048, and 0.928 for the fake trajectory. 
The proposed method can also perform black box attack on different tracking architectures.
We will make our code publicly available.
\end{abstract}

% 关键字
\begin{IEEEkeywords}
Visual tracking, adversarial attacks, Siamese networks.
\end{IEEEkeywords}
\IEEEpeerreviewmaketitle

@@

Recently, the robustness of Siamese trackers have attracted much attention in the sense of testing their vulnerability to adversarial attacks, and the focus has been directed towards more efficient and low-cost attacks \cite{TTP,FAN,SPARK}. However, these attack methods still need to craft the perturbations for each video independently based on either iterative optimisation or adversarial network inference, for which the adequacy of computational resources may be not ensured by the computationally intensive tracking systems.

Universal adversarial perturbations (UAPs) proposed in \cite{UAP} can fool most images from a data distribution in an image-agnostic manner. Being universal, UAPs can be conveniently exploited to perturb unseen data on-the-fly without extra computations. However, no existing work has touched the topic of attacking the Siamese trackers using UAPs, because it is hard to apply existing UAPs to attack Siamese trackers directly. The main reason lies in the fact that, (a) most UAPs are designed for typical neural networks with one image as input while Siamese networks accept both the template and search images, and (b) the goal of existing UAP methods is to disturb unary or binary model outputs for single instance while we need to use universal perturbations to mislead Siamese trackers to follow a specified trajectory.

@@

To sum up, the main contributions of this paper are:
\begin{itemize}
  \item To the best of our knowledge, we are the first to perform universal targeted attacks for Siamese trackers utilizing both the imperceptible perturbation and the adversarial patch together, which are jointly trained in an end-to-end manner.
  \item We demonstrate that the adversarial patch can be rather small: the 16-by-16 adversarial patch can make the AO score drop dramatically to 0.12 and its area only accounts for 0.39\% of the search image area.
  \item Although trained using the parameters of the SiamFC++ network, the universal perturbations can perform black box attack on other trackers such as SiamRPN and SiamRPN++.
\end{itemize}

@@

\textit{White box attacks v.s. Black box attacks} In the white box attack setting \cite{meng2019white}, the adversary has full knowledge of the model including model type, model architecture and values of all parameters and trainable weights. In the black box setting \cite{cheng2018query,li2019nattack,papernot2017practical,li2020projection}, the adversary has limited or no knowledge about the model under attack \cite{kurakin2018adversarial}. In this paper, we focus on the white box attack for Siamese trackers. However, although we train perturbations using the parameters of the SiamFC++ network, we surprisely find that the trained perturbations can perform black attack on other trackers such as SiamRPN and SiamRPN++.

\textit{Non-universal attacks v.s. Universal attacks} In the setting of non-universal attacks \cite{dai2018adversarial,li2018second,lin2017tactics}, the adversary has to generate one perturbation of every new datapoint, which is time consuming. In the setting of the universal attacks \cite{khrulkov2018art,mopuri2018nag,zhang2020understanding,mopuri2018generalizable,chen2018shapeshifter}, only one perturbation is generated and used for every datapoint of the dataset. In this paper, we focus on generating universal perturbations for Siamese trackers to perform efficient attack.

@@

Let $V=\{I_i\}_1^T$ denote the frames of a video sequence of length $T$. $B^{gt}=\{b^{gt}_i\}_1^T$ is used to represent the target's ground-truth position in each frame. The visual object tracking aims to predict the position $B^{pred}=\{b^{pred}_i\}_1^T$ of this target in the subsequent frames given its initial state. In SiamFC++, the tracker first transforms the paired reference frame $I_1$ and annotation $b_1^{gt}$ to get an template image $\textbf z$, and transforms the search frame $I_i$ to get the search image $\textbf x_i$ centered at the position estimated in the previous frame. At each time-step, the template image $\textbf z$ and the search image $\textbf x_i$ are first passed individually through a shared backbone network and then fused using a channel-weise correlation operation:
\begin{equation}
  f_{i}(\mathbf{z}, \mathbf{x})=\psi_{i}(\phi(\mathbf{z})) \star \psi_{i}(\phi(\mathbf{x})), i \in\{\mathrm{cls}, \text {reg}\}
\end{equation}
where $\star$ denotes the channel-wise correlation operation, $\phi(\cdot)$ denotes the feature extractor of the Siamese network, $\psi_i(\cdot)$ denotes the layer specific to the classification or regression task, and $i$ denotes the specific task type ($\mathrm{cls}$ denotes the classification task and $\mathrm{reg}$ denotes the regression task). $\psi_{\mathrm{cls}}$ and $\psi_{\mathrm{reg}}$ are both designed as two convolutional layers for adapting generic features to the feature space specific to the classification/regression task. The fused features then act as input to a head network, which predicts a classification map $\textbf{C}$, a bounding box regression map $\textbf{R}$, and a quality assessment map $\textbf{Q}$ in an anchor-free manner. In short, $\textbf C$ encodes the probability of each spatial position to contain the target, $\textbf R$ regresses the bounding box of the target, and $\textbf Q$ predicts the target state estimation quality. The final bounding box is then generated according to $\textbf{C}$, $\textbf{R}$ and $\textbf{Q}$.

@@

Our attack algorithm is based on the imperceptible perturbation genration method \cite{FGSM} and adversarial patch generation method \cite{patch}. However, we do not simply apply either of above methods to trackers. In this paper, we demostrate that the imperceptible perturbations and the adversarial patch are indispensable for attacking trackers and can be jointly optimized in and end-to-end training manner. On the one hand, the small patch applied to the search image works as a fake target, which tells the tracker where to look. It is crucial to design a strategy for the tracker to follow a predefined trajectory. For example, the TTP method \cite{TTP} pre-computes perturbations corresponding diverse directions and use them to force the tracker to follow the target trajectory. However, this method is sub-optimal because the tracker only predicts the approximation of the specified trajectory. As shown in experiments, the targeted attack performance of TTP is not as good as ours. On the other hand, the imperceptible perturbation applied to the template image is optimized to find the fake target, cooling down hot regions where the real target exists on the heatmaps and increasing the responses at the position of the fake target (see Figure \ref{fig:1}).

@@

In this subsection, we show how to train the video-agnostic perturbations $(\delta, p)$ for Siamese trackers (Figure \ref{fig:vis_perturbations}). At beginning, each element in $\delta$ is initialized to 0. $p$ is initialized with a normal distribution, with a mean of 127 and a standard deviation of 1. During the $k$-th iteration of training, a video $V=\{I_i\}_1^T$ is randomly selected from the training dataset $\mathcal V$. Assuming the template perturbation at the $k$-th iteration is $\delta_k \in \mathbb{R}^{127\times 127 \times 3}$, and the adversarial patch is $p_k \in \mathbb{R}^{128\times 128\times 3}$. We first randomly pick paired frames $I_t, I_s$ from $V$. The clean template image $\textbf z\in \mathbb{R}^{127\times 127 \times 3}$ is generated according to $I_t$ and $b^{gt}_t$, and the perturbed template image is:
\begin{equation}
\tilde {\textbf z} = \textbf z + \delta_k.
\end{equation}
Similarly, the clean search image $\textbf x \in \mathbb{R}^{303\times 303 \times 3}$ is generated according to $I_s$ and $b^{gt}_s$.
As mentioned before, the patch is regarded as a \textit{fake target} and pasted into the search images. We force the center position of the \textit{fake target} to near the center position of the real target within a shift range of 64 pixels, where shift is defined as the maximum range of translation generated from a uniform distribution.
The width/height of the \textit{fake target} is randomly selected between 32 pixels and 128 pixels.
The perturbed search image is generated as follows:
\begin{equation}
\tilde{\textbf x} = A(\textbf x, p_k, (x_0, y_0, x_1, y_1)),
\end{equation}
where $(x_0, y_0)$ and $(x_1, y_1)$ denote the coordinates of the upper-left and lower-right corners of the fake target on the search image. $A$ is a patch application operator \cite{patch} which first resizes the patch $p_k \in \mathbb{R}^{128\times 128\times 3}$ to $\hat{p}_k \in \mathbb{R}^{w\times h\times 3}$, and then pastes the resized patch $\hat{p}_k$ into the search image $\textbf x$ at location $(l^x,l^y)$. Subsequently, the SiamFC++ tracker takes $\tilde {\textbf x}$ and $\tilde{\textbf  z}$ as input and predicts $\textbf{C, R, Q}$ in an anchor-free manner.

\begin{algorithm}[tb]
  \caption{Training Process}
  \label{alg:algorithm}
  \textbf{Input}: Training dataset $\mathcal{V}$, Siamese tracker $\phi$, and max iteration number $N$.\\
  \textbf{Output}: Imperceptible perturbation $\delta$, and adversarial patch $p$.
  \begin{algorithmic}[1] %[1] enables line numbers
  \STATE Let $k = 0$.
  \WHILE{$k < N$}
  \STATE Randomly pick a video $V\in \mathcal{V}$. The corresponding ground truth is $B^{gt}=\{b^{gt}_i\}^T_1$.
  \STATE Randomly pick paired frames $I_t, I_s$ from $V$.
  \STATE Generate template image $\textbf{z}$ according to $I_t$ and $b^{gt}_t$.
  \STATE $\tilde{\textbf{z}} = \textbf{z} + \delta_k.$
  \STATE Generate search image $\textbf{x}$ according to $I_s$ and $b^{gt}_s$.
  \STATE Calculate the \textit{fake target} position $\{x_0, y_0, x_1, y_1\}$ with respect to the search image.
  \STATE $\tilde{\textbf x} = A(\textbf x, p_k, (x_0, y_0, x_1, y_1)).$
  \STATE $\textbf{C, R, Q} = \phi(\tilde {\textbf x}, \tilde{\textbf z}).$
  \STATE Generate fake labels $\textbf{C}^*,\textbf{R}^*,\textbf{Q}^*$ using $\{x_0, y_0, x_1, y_1\}$.
  \STATE Calculate loss $L(\textbf{C, R, Q}, \textbf{C}^*, \textbf{R}^*, \textbf{Q}^*)$ using Equ. \ref{eq:loss}.
  \STATE $\delta_{k+1} = \delta_{k} - \epsilon_1 \cdot \text{sign}(\nabla_{\delta_k}L).$
  \STATE $p_{k+1} = p_{k} - \epsilon_2 \cdot \text{sign}(\nabla_{p_k}L).$
  \STATE $k = k + 1.$
  \ENDWHILE
  \STATE \textbf{return} $\delta_N, p_N.$
  \end{algorithmic}
  \label{alg}
\end{algorithm}

\textit{Generating Fake Labels} The fake lables are composes of three parts: fake classification label $\textbf{C}^*$, fake regression label $\textbf{R}^*$ and fake quality estimation label $\textbf{Q}^*$.

For classification, location $(x,y)$ on feature map $\psi_{\mathrm{cls}}$ is considered as a positive sample (i.e., $\textbf{C}^*_{x,y} = 1$) if its corresponding location $\left(\left\lfloor\frac{s}{2}\right\rfloor+x s,\left\lfloor\frac{s}{2}\right\rfloor+y s\right)$ on the input image falls into the fake bounding box, and vice versa are considered as negative samples (i.e., $\textbf{C}^*_{x,y} = 0$). $s=8$ denotes the total step length of the feature extraction network.

In the regression branch, the last convolution layer predicts the distance of position $\left(\left\lfloor\frac{s}{2}\right\rfloor+x s,\left\lfloor\frac{s}{2}\right\rfloor+y s\right)$ to the four edges of the fake ground truth bounding box, denoted as the four-dimensional vector $\boldsymbol{t}^{*}=\left(l^{*}, t^{*}, r^{*}, b^{*}\right)$. Thus, the fake regression label $\textbf{R}^*$ at position $(x,y)$ can be expressed as:

\begin{equation}
  \begin{array}{ll}
  l^{*}=\left(\left\lfloor\frac{s}{2}\right\rfloor+x s\right)-x_{0}, \quad t^{*}=\left(\left\lfloor\frac{s}{2}\right\rfloor+y s\right)-y_{0} \\
  r^{*}=x_{1}-\left(\left\lfloor\frac{s}{2}\right\rfloor+x s\right), \quad b^{*}=y_{1}-\left(\left\lfloor\frac{s}{2}\right\rfloor+y s\right)
  \end{array}
\end{equation}

SiamFC++ assumes that feature pixels near the center of the target will have higher importance than other pixels. Following the design of SiamFC++, we use a $1 \times 1$ convolutional layer for quality assessment, i.e., to learn the intersection over union (IoU) score of the predicted box and $B^{fake}$:

\begin{equation}
  \mathrm{IoU}^{*}=\frac{\text { Intersection }\left(B^{pred}, B^{fake}\right)}{\operatorname{Union}\left(B^{pred}, B^{fake}\right)}
\end{equation}

@@

\textit{Ablation Study: Influence of Training Iterations} Table \ref{tab:iter} and Figure \ref{fig:iter} demonstrate the influence of the number of training iterations on the attack performance. It can be observed that as the number of iterations increases, the AO score with respect to the fake trajectory increases significantly, while the AO score with respect to the real trajectory decreases significantly. This demonstrates the effectiveness of the end-to-end training process proposed in this paper. After about 30,000 training iterations, the resulting perturbation prevents the tracker from tracking most of the targets in GOT-Val, and the AO with respect to the real trajectory decreases from 0.760 to 0.035. Note that the AO decreases significantly faster at the beginning (training iterations less than 2048). This demonstrates the fast convergence capability of our end-to-end training scheme.

To verify whether the perturbations generated for the template images are difficult to perceive, we evaluate the quality of the template images after the perturbations. The quality of the perturbed template images is evaluated by the metric SSIM. As shown in Table \ref{tab:iter}, the SSIM value of the perturbed template image gradually decreases as the training proceeds, which indicates that the difference between the perturbed template image and the original template image gradually increases due to the learned adversarial perturbation being added to the clean template image. The SSIM value ranges from 0 to 1. If the SSIM is close to 1, the difference between the perturbed template image and the original template image is small. If SSIM is closer to 0, it means that the difference between the perturbed template image and the original template image is large, which means that the perturbation generated by the template image is easier to perceive. The SSIM value after convergence of this algorithm is 0.88, which means that the difference between the perturbed template image and the original template image is small, that is, the perturbation generated by the template image is harder to perceive (see Figure \ref{fig:1}).

@@

% 作者
%\vfill  % 距离更远了
%\enlargethispage{-5in}  % 距离近了些
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{images/zhbli.jpg}}]
\begin{IEEEbiographynophoto}
{Zhenbang Li}
received the B.S. degree in computer science and technology from Beijing Institute of Technology, Beijing, China, in 2016. Currently, he is a Ph.D. student in Institute of Automation, Chinese Academy of Sciences, Beijing, China. His research interests include object tracking and deep learning.
\end{IEEEbiographynophoto}

\begin{IEEEbiographynophoto}
{Jin Gao}
received the B.S. degree from the Beihang University, Beijing, China in 2010, and the Ph.D. degree from the University of Chinese Academy of Sciences (UCAS) in 2015. Now he is an assistant professor with the National Laboratory of Pattern Recognition, Institute of Automation, Chinese Academy of Sciences (CASIA). His research interests include visual tracking, autonomous vehicles, and service robots. 
\end{IEEEbiographynophoto}

\begin{IEEEbiographynophoto}
{YaYa Shi}
received the B.S. degree from the Hefei University of Technology, Anhui, China in 2018. Currently, he is a Ph.D. student in the University of Science and Technology of China, Anhui, China. Her research interests include video captioning and deep learning.
\end{IEEEbiographynophoto}

\begin{IEEEbiographynophoto}
{Bing Li}
received the Ph.D. degree from the Department of Computer Science and Engineering, Beijing Jiaotong University, China, in 2009. He is currently a Professor in the Institute of Automation, Chinese Academy of Sciences, Beijing, China. His research interests include video understanding, color constancy, visual saliency, and web content mining.
\end{IEEEbiographynophoto}
